Online (Bandit) Policy Selection for EASY-Backfilling

### 2018-01-07 paper/src/paper.tex
modifs

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards\cite{bubnow}. At every round, an action (represented by an arm
of the bandit) must be chosen from a fixed set and the corresponding reward is
observed. The goal of a MAB algorithm is to maximize the total reward obtained
in a successive number of rounds. There exist multiple works addressing this
problem under a variety of constraints, The two most popular settings are the
original stochastic  case and the adversarial case. See~\cite{thompson} for the
original work on the stochastic case and~\cite{Auer2002} for the ''Upper
Confidence Bound`` family of algorithms. To the best of our knowledge, the work
of~\cite{Banos} is the earliest for the adversarial case; see~\cite{nonstoch}
for the ``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow}
for a comprehensive review of the field. While these algorithms bound the
cumulative difference in loss to the best arm (the regret), they have
functional constraints such as the fact that the rewards should be contained in
a range. The simple heuristic called \textit{Epsilon-Greedy} introduced
in~\cite{Auer2002} does not have this requirement.  This heuristic will be used
in this paper and is described in Subsection~\ref{sub:bandit}

@@

The EASY heuristic uses a job queue to select and backfill jobs. While this job
queue is ordered in FCFS order in the original heuristic, it is possible to
reorder it at will. We choose on a search space of 10 reordering policies.

@@

\begin{algorithm}[]
  \caption{EASY($P,\Theta$) policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE Queue $Q$ of waiting jobs.
    \ENSURE None (calls to $Start()$)
    \STATE Sort $Q$ according to $P$
    \STATE Move all jobs of $Q$ for which $wait_j > \Theta$ ahead
    of the queue (breaking ties in FCFS order).
    \\ \textit{//Starting jobs until the machine is full}
    \FOR {job $j$ in Q}
    \IF {$j$ can be started given the current system use}
    \STATE Pop $j$ from Q
    \STATE $Start(j)$
    \ELSE
    \STATE Reserve $j$ at the earliest
    time possible according to the estimated running times
    of the currently running jobs.
    \\ \textit{//Backfilling jobs}
    \FOR {job $j'$ in $Q\setminus\{j\}$}
    \IF {$j'$ can be started without delaying $j$'s reservation.}
    \STATE Pop $j'$ from $Q$
    \STATE $Start(j')$
    \ENDIF
    \ENDFOR
    \STATE \textbf{break}
    \ENDIF
    \ENDFOR
  \end{algorithmic}
  \label{alg:EASY}
\end{algorithm}

@@

Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during
the period $\Delta_t$. We define the cost of policy $P_i$ during period
$\Delta_t$ as:
%
\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta_t)}
  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t
  \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
%
As one can note, the actual waiting time is used here, so that one dispenses
with the use of a simulator for efficiency considerations. This leads to a
faster and easier estimate of the cumulative waiting time, however only for the
policy that is used during $\Delta_t$. Note that there is as previously a bias
due to boundary effects in the measurement method, as we do include jobs that
had been started before $\Delta_t$, as well as disregard jobs that were
submitted during $\Delta_t$ but finish later.

%This also explains the difference
%between the definitions of $l(\Delta_t)$: in the previous selection strategies,
%one relies on a simulated estimate so that jobs need not be finished during
%$\Delta_t$, but only submitted; in contrast, to know their actual waiting time,
%jobs need be finished during $\Delta_t$.

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
\frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
$\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary to ensure that policies
remain comparable over time. Not normalizing would favor the policy that was
first selected, even if this choice was random. This normalization however
entails that the size of the policy period is important: it should not be too
large so as to avoid relying on too few points for estimating the cost, and it
should not be too small either to avoid extreme variations between periods. As
explained in Section~\ref{sec:experiments}, we rely in this study on periods of
one day and one week.

@@

  \item If the number of allocated or requested cores is negative, we use the
    available positive value as the request. If both are negative, we remove
    the job;

  \item If the runtime or submission time is negative, we remove the job.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} suggests to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled.  This simple mechanism
enables us to preserve the dependency within a week.  Although the system state
may not be independent from one week to an other, we rely on existing
work~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. While we do provide guidelines as how to incorporate tracking
variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$, this will not be studied in the experiments where 
$\lambda=1$, and is left for future research.

@@

We publish a declarative workflow that runs simulations and generates all the
figures from this paper. Any question regarding replication can be freely
directed to the authors via e-mail.

There are various approaches for ensuring computational experiments are replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach\ref{nix}. This allows to replicate the experiments on
clusters where virtualization is not available or kernels are too old for
containers and decreases simulation runtime.

@@

All the code associated with this work is released under the~\cite{isc}
license.
Running the experiments can be done
on any platform equipped with the Nix package manager by running at the root of
the extracted archive:

@@

 \begin{itemize}
   \item Data. Data extraction from archives, filtering, resampling.
     this is principally using shell scripts for data manipulation (see file
     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong_filter| for the
     filtering steps used) and ocaml code that implements the resampling method
     described in~\ref{ssub:resampling}.

   \item Simulation. This step runs the lightweight ocaml backfilling
     simulator specially written for this work. this simulator is made
     available under the isc~\cite{isc} license as a persistent zenodo
     archive at~\cite{ocst}.

@@

  \caption{Follow-up from figure~\ref{fig:small}, plot for the UniLu-Gaia log.}

@@

Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
results concerning the strategies used in this paper. These figures show the
average cumulative improvement of over EASY(FCFS,40h) of the three strategies
we use in the same format as Figure~\ref{fig:all}. There are five
strategies displayed on the graph:

@@

\begin{table*}[]
  \centering
  \ra{1.3}
  \caption{Average Cumulative waiting time improvement of the policies used with respect to EASY(FCFS).}
  \label{tab:strat}
  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
    \hline
    Trace        & Random &       & Full           &                & Noisy &       & Bandit & \\
    \hline
    $\Delta$     & week   & day   & week           & day            & week  & day   & week   & day \\
    KTH-SP2      & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
    CTC-SP2      & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
    SDSC-SP2     & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
    SDSC-BLUE    & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
    ANL-Intrepid & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
    CEA-Curie    & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
    Unilu-Gaia   & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
    \hline
  \end{tabular}
\end{table*}

@@

\textbf{FCFS vs Random.} Randomly taking a policy in the search set defined in
~\ref{subsec:policies} is better than using the FCFS order. This is due to the
known fact that the FCFS order is not optimal for optimizing the average
waiting time objective, and confirms that the search set is well chosen.  The
value of the period length $\Delta$ has an impact on the performance of the
random policy. Indeed, shorter periods lead to better performance. This is perhaps
due to the use of queuing in the scheduling process:
When a bad policy is used, the job queue builds up. By switching more
frequently from bad to good policies, the system never has time to build up the
job queue. In other words, while bad decisions impact the system through an
accumulation of waiting jobs, there is less contribution to system inertia from
good decisions, as the number of jobs in queue can not be negative. This
interpretation also means that randomized bandit strategies are robust by
design, and needs more subsequent experimenting to be validated.

>>>>

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards\cite{bubnow}. At every round, an action called an
"arm" in the litterature must be chosen from a fixed set and the corresponding
reward is observed. The goal of a MAB algorithm is to maximize the total reward
obtained in a successive number of rounds. This is essentially achieved using a
combination of explorative actions, which help estimate the quality of each
arm, and of exploitative actions, which leverage the current estimates by using
a good arm. This problem can be addressed using various hypotheses. Most common
are the (original) stochastic setting and the adversarial setting of regret
minimization. See~\cite{thompson} for the original work on the stochastic
setting and~\cite{Auer2002} for the ''Upper Confidence Bound`` family of
algorithms. To the best of our knowledge, the work of~\cite{Banos} is the
earliest for the adversarial setting; see~\cite{nonstoch} for the
``Exponential-weight'' family of algorithms. We refer to~\cite{bubnow} for a
comprehensive review of the field. While these algorithms bound the cumulative
difference in loss to the best arm (the regret), they have functional
constraints such as the fact that the rewards should be contained in a range.
The simple heuristic called \textit{Epsilon-Greedy} introduced
in~\cite{Auer2002} does not have this requirement. For this reason, this
heuristic will be used in this paper and is described in
Subsection~\ref{sub:bandit}.

@@

The EASY heuristic uses a job queue to select and backfill jobs. While this job
queue is ordered in FCFS order in the original heuristic, it is possible to
reorder it at will. We focus on a search space of 10 reordering policies.

@@

\begin{algorithm}[]
  \caption{EASY($P,\Theta$) policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE Queue $Q$ of waiting jobs.
    \ENSURE None (calls to $Start()$)
    \STATE Sort $Q$ according to $P$
    \STATE Move all jobs of $Q$ for which $wait_j > \Theta$ ahead
    of the queue (breaking ties in FCFS order).
    \\ \textit{//Starting jobs until the machine is full}
    \FOR {job $j$ in Q}
    \IF {$j$ can be started given the current system use}
    \STATE Pop $j$ from Q
    \STATE $Start(j)$
    \ELSE
    \STATE Reserve $j$ at the earliest
    time possible according to the estimated running times
    of the currently running jobs.
    \\ \textit{//Backfilling jobs}
    \FOR {job $j'$ in $Q\setminus\{j\}$}
    \IF {$j'$ can be started without delaying the reservation on $j$.}
    \STATE Pop $j'$ from $Q$
    \STATE $Start(j')$
    \ENDIF
    \ENDFOR
    \STATE \textbf{break}
    \ENDIF
    \ENDFOR
  \end{algorithmic}
  \label{alg:EASY}
\end{algorithm}

@@

Let $l(\Delta^{'}_t)$ denote this time the number of jobs \textit{finished} during
the period $\Delta^{'}_t$. We define the cost of policy $P_i$ during period
$\Delta^{'}_t$ as:
%
\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta^{'}_t)}
  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta^{'}_t
  \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
%
As one can note, the actual waiting time is used here, so that one dispenses
with the use of a simulator for efficiency considerations. This leads to a
faster and easier estimate of the cumulative waiting time, however only for the
policy that is used during $\Delta^{'}_t$. Note that there is as previously a bias
due to boundary effects in the measurement method, as we do include jobs that
had been started before $\Delta^{'}_t$, as well as disregard jobs that were
submitted during $\Delta^{'}_t$ but finish later.

%This also explains the difference
%between the definitions of $l(\Delta^{'}_t)$: in the previous selection strategies,
%one relies on a simulated estimate so that jobs need not be finished during
%$\Delta^{'}_t$, but only submitted; in contrast, to know their actual waiting time,
%jobs need be finished during $\Delta^{'}_t$.

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation} \label{eq:tot-cost-greedy} w(\rightarrow T;P_i) =
  \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta^{'}_t)} \sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i) \end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period
$\Delta^{'}_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1}
\mathbbm{1}(P_i,t) l(\Delta^{'}_t)$ is necessary to ensure that policies
remain comparable over time. Not normalizing would favor the policy that was
first selected, even if this choice was random. This normalization however
entails that the size of the policy period is important: it should not be too
large so as to avoid relying on too few points for estimating the cost, and it
should not be too small either to avoid extreme variations between periods. As
explained in Section~\ref{sec:experiments}, we rely in this study on periods of
one day and one week.

@@

  \item If the number of allocated or requested cores is negative, we use the
    available positive value as the request. If both are negative (no data is
    available), we remove the job;

  \item If the runtime or submission time is negative (no data is available),
    we remove the job.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
lose the details of the workloads. This is problematic due to the non-linear
aspect of the job scheduling process. This is discussed
in~\cite{feitresampling}, an approach which attempts to solve this problem by
resampling data directly from the original traces. In this work, we use a
simple implementation of these ideas. More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled. This simple mechanism enables
us to preserve the dependency within a week. Although the system state may not
be independent from one week to an other, we rely on existing
work~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. While we do provide guidelines as how to incorporate tracking
variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$, this will not be studied in the experiments where
$\lambda=1$, and is left for future research.

@@

We publish a declarative workflow~\cite{repro} based on our
simulator~\cite{ocst} that runs experiments and generates all the figures from
this paper.

There are various approaches for ensuring computational experiments are replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach~\cite{nix}. This allows to replicate the experiments on
clusters where virtualization is not available or kernels are too old for
containers and decreases simulation runtime.

@@

All the code associated with this work is released under the ISC~\cite{isc}
license.
Running the experiments can be done
on any platform equipped with the Nix package manager by running at the root of
the extracted archive:

@@

 \begin{itemize}
   \item Data. Data extraction from archives, filtering, resampling.
     This is principally using shell scripts for data manipulation (see file
     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong_filter| for the
     filtering steps used) and ocaml code that implements the resampling method
     described in~\ref{ssub:resampling}.

   \item Simulation. This step runs the lightweight ocaml backfilling
     simulator specially written for this work. this simulator is made
     available under the ISC~\cite{isc} license as a persistent zenodo
     archive at~\cite{ocst}.

@@

  \caption{Figure~\ref{fig:small} continued, plot for the UniLu-Gaia log.}

@@

Figures~\ref{fig:small} and~\ref{fig:follow} report the main experimental
results concerning the strategies used in this paper. These figures show the
average cumulative improvement over EASY(FCFS,40h) of the three strategies
we use in the same format as Figure~\ref{fig:all}. There are five
strategies displayed on the graph:

@@

\begin{table*}[]
  \centering
  \ra{1.3}
  \caption{Average Cumulative waiting time improvement of the policies used with respect to EASY(FCFS).}
  \label{tab:strat}
  \begin{tabular}{@{}lrrrrrrrrrrrr@{}}
    \hline
    Trace        & Best(hindsight) & Random &       & Full           &                & Noisy &       & Bandit & \\
    \hline                
    $\Delta$     &        & week   & day   & week           & day            & week  & day   & week   & day \\
    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
    CTC-SP2      & -47\%  & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
    SDSC-BLUE    & -32\%  & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
    ANL-Intrepid & -28\%  & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
    Unilu-Gaia   & -62\%  & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
    \hline
  \end{tabular}
\end{table*}

@@

\textbf{FCFS vs Random.} Randomly taking a policy in the search set defined in
~\ref{subsec:policies} is better than using the FCFS order. This is due to the
known fact that the FCFS order is not optimal for optimizing the average
waiting time objective, and confirms that the search set is well chosen. The
value of the period length $\Delta$ has an impact on the performance of the
random policy. Indeed, shorter periods lead to better performance.
